The Impact of Interview Disruptions on Data Quality in Social Surveys

### 2024-12-03 paper/paper.qmd
update paper

In recent years, the quality and reliability of data collected through social surveys have garnered increasing attention, especially considering how various forms of disruption may influence respondents’ behavior and, subsequently, the quality of survey responses. Social surveys are indispensable tools for obtaining observations into societal trends, guiding public policy, and informing evidence-based decision-making. Thus, ensuring the robustness of data collected from these surveys has implications for the accuracy of the conclusions drawn and for the effective formulation of policy measures. Previous research has made significant strides in understanding factors that impact survey data quality. For example, Groves et al. [@groves2011survey]investigated the influence of interviewer characteristics—such as demeanor, experience, and communication style—on the quality of data obtained, employing a multi-level modeling approach to assess these refined dynamics. Similarly, Couper and de Leeuw [@massey2013we]explored how the survey mode (e.g., in-person versus telephone) affects respondent comfort and attentiveness, highlighting that face-to-face interviews often elicit more thorough responses due to the availability of visual cues. In another significant contribution, Fowler and Mangione [@aquilino1993effects]analyzed the impact of third-party presence during interviews, finding that interruptions can lead to response biases, thereby compromising data integrity. Despite these advances, there remains a clear gap that needs to be filled: the distinct effects of different types of disruptions—specifically those involving family members or occurring due to environmental factors—on respondent comprehension and the overall quality of responses have not yet been systematically explored. Addressing this gap is paramount to improving the reliability of survey data, particularly in settings where such disruptions are unavoidable.

@@

# Data {#sec-data}
## Data Source
The dataset for this study is sourced from the [European Social Survey](https://ess.sikt.no/en/datafile/0c3f5fda-4fe0-42f2-b4c8-4ef093c0508f/7?tab=0&elems=de8c08b6-e406-4c56-92ee-4272818cae96_1)[@ess_datafile], a widely recognized repository that provides detailed information on societal attitudes and behaviors across numerous European countries. The dataset includes various subsets designed for distinct stages of analysis, enabling an examination of how factors like disruptions influence survey responses. The primary dataset used for this research contains 9,743 rows and 25 variables, capturing details such as types of disruptions (e.g., partner or child interference), methods of interview, and characteristics of respondents. The standardized collection procedures employed by the ESS ensure consistency across countries, making the dataset well-suited for cross-national comparisons.

## Overview
We utilized the statistical programming language R [@citeR]. to prepare, analyze, and visualize the survey data. This study focuses on understanding how interview disruptions affect response quality by examining variables related to both interviewer and respondent characteristics. The structured design of the dataset allowed for a thorough exploration of contextual disruptions and their influence on data reliability, providing valuable insights for enhancing survey methodologies.

@@

We ran the model in R [@citeR] using the `rstanarm` package@rstanarm. The default priors from `rstanarm` provide an appropriate level of smoothing and regularization, making the model applicable to survey data with complex patterns while still allowing the data to reveal meaningful relationships, especially concerning factors that affect response quality.

>>>>

In recent years, the quality and reliability of data collected through social surveys have garnered increasing attention, especially considering how various forms of disruption may influence respondents’ behavior and, subsequently, the quality of survey responses. Social surveys are indispensable tools for obtaining observations into societal trends, guiding public policy, and informing evidence-based decision-making. Thus, ensuring the robustness of data collected from these surveys has implications for the accuracy of the conclusions drawn and for the effective formulation of policy measures. Previous research has made significant strides in understanding factors that impact survey data quality. For example, Groves et al [@groves2011survey]. investigated the influence of interviewer characteristics—such as demeanor, experience, and communication style—on the quality of data obtained, employing a multi-level modeling approach to assess these refined dynamics. Similarly, Couper and de Leeuw [@massey2013we]. explored how the survey mode (e.g., in-person versus telephone) affects respondent comfort and attentiveness, highlighting that face-to-face interviews often elicit more thorough responses due to the availability of visual cues. In another significant contribution, Fowler and Mangione [@aquilino1993effects].analyzed the impact of third-party presence during interviews, finding that interruptions can lead to response biases, thereby compromising data integrity. Despite these advances, there remains a clear gap that needs to be filled: the distinct effects of different types of disruptions—specifically those involving family members or occurring due to environmental factors—on respondent comprehension and the overall quality of responses have not yet been systematically explored. Addressing this gap is paramount to improving the reliability of survey data, particularly in settings where such disruptions are unavoidable.

@@

# Data {#sec-data}
## Data Source
The dataset for this study is sourced from the [European Social Survey](https://ess.sikt.no/en/datafile/0c3f5fda-4fe0-42f2-b4c8-4ef093c0508f/7?tab=0&elems=de8c08b6-e406-4c56-92ee-4272818cae96_1)[@ess_datafile]., a widely recognized repository that provides detailed information on societal attitudes and behaviors across numerous European countries. The dataset includes various subsets designed for distinct stages of analysis, enabling an examination of how factors like disruptions influence survey responses. The primary dataset used for this research contains 9,743 rows and 25 variables, capturing details such as types of disruptions (e.g., partner or child interference), methods of interview, and characteristics of respondents. The standardized collection procedures employed by the ESS ensure consistency across countries, making the dataset well-suited for cross-national comparisons.

## Overview
We utilized the statistical programming language R [@citeR] to prepare, analyze, and visualize the survey data. This study focuses on understanding how interview disruptions affect response quality by examining variables related to both interviewer and respondent characteristics. The structured design of the dataset allowed for a thorough exploration of contextual disruptions and their influence on data reliability, providing valuable insights for enhancing survey methodologies.

@@

We ran the model in R [@citeR] using the `rstanarm` package @rstanarm. The default priors from `rstanarm` provide an appropriate level of smoothing and regularization, making the model applicable to survey data with complex patterns while still allowing the data to reveal meaningful relationships, especially concerning factors that affect response quality.
